%Copyright 2019 Christopher M. Jermaine (cmj4@rice.edu) and Risa B. Myers (rbm2@rice.edu)
%
%Licensed under the Apache License, Version 2.0 (the "License");
%you may not use this file except in compliance with the License.
%You may obtain a copy of the License at
%
%    https://www.apache.org/licenses/LICENSE-2.0
%
%Unless required by applicable law or agreed to in writing, software
%distributed under the License is distributed on an "AS IS" BASIS,
%WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
%See the License for the specific language governing permissions and
%limitations under the License.
%===============================================================
\documentclass[aspectratio=169]{beamer}

\mode<presentation> 
{
\usetheme[noshadow, minimal,numbers,riceb,nonav]{Rice}
\usefonttheme[onlymath]{serif}
\setbeamercovered{transparent}
}
\useinnertheme{rectangles}

\usepackage[english]{babel}

\usepackage{amsmath}
\usepackage{mathptmx}
\usepackage{helvet}
\usepackage{courier}
\usepackage[T1]{fontenc}
\usepackage{trajan}
\usepackage{ textcomp }
\renewcommand{\footnotesize}{\tiny}


\usepackage{listings}

\newenvironment{noindentitemize}
{ \begin{itemize}
 \setlength{\itemsep}{1.5ex}
  \setlength{\parsep}{0pt}   
  \setlength{\parskip}{0pt}
 \addtolength{\leftskip}{-2em}
 }
{ \end{itemize} }

\newenvironment{noindentitemize2}
{ \begin{itemize}
  \setlength{\itemsep}{0ex}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}   
  \addtolength{\leftskip}{-2em}  }
{ \end{itemize} }

\lstnewenvironment{SQL}
  {\lstset{
        aboveskip=5pt,
        belowskip=5pt,
        escapechar=!,
        mathescape=true,
        upquote=true,
        language=C,
        basicstyle=\linespread{0.94}\ttfamily\normalsize,
        deletekeywords={VALUE, PRIOR},
        showstringspaces=true}
        \vspace{0pt}%
        \noindent\minipage{0.47\textwidth}}
  {\endminipage\vspace{0pt}}
  
\newcommand{\LIKES}{\textrm{LIKES}} 
\newcommand{\FREQUENTS}{\textrm{FREQUENTS}} 
\newcommand{\SERVES}{\textrm{SERVES}} 
\newcommand{\CAFE}{\textrm{CAFE}} 
\newcommand{\COFFEE}{\textrm{COFFEE}} 
\newcommand{\DRINKER}{\textrm{DRINKER}} 
\newcommand{\ALLPEEPS}{\textrm{ALLPEEPS}} 
\newcommand{\ALLCOMBOS}{\textrm{ALLCOMBOS}} 

\setbeamerfont{block body}{size=\tiny}

%===============================================================%

\title[]
{Tools \& Models for Data Science}

\subtitle{Course Wrap-up}

\author[]{Chris Jermaine \& Risa Myers}
\institute
{
  Rice University 
}

\date[]{}

\subject{Beamer}

\begin{document}

\begin{frame}
 \titlepage
\end{frame}
%***********************************************************
\begin{frame}{Course Objectives}

At the end of this course, students will understand the \textbf{development and use of modern machine learning tools}, including Spark and will be able to \textbf{implement machine learning algorithms} using this system. They will have basic skills in \textbf{querying relational databases} and will understand and be able to \textbf{implement and use common data science models}, including gradient descent, K-nearest neighbors, and more. They will also \textbf{be familiar with the theoretical basis and underlying research that motivated the systems and models discussed in class}.

\end{frame}
%***********************************************************
\begin{frame}{What have we covered?}

\begin{itemize}
\item Databases
\item Systems \& Tools
\item Models
\item Concepts
\end{itemize}
\end{frame}
%***********************************************************
\begin{frame}{Databases}

\begin{itemize}
\item The Relational Model
\item Relational Calculus
\item SQL
\end{itemize}
\end{frame}
%***********************************************************
\begin{frame}{Systems \& Tools}

\begin{itemize}
\item Jupyter Notebooks
\item Command line tools
\item MapReduce
\item Spark
\item AWS
\item Numpy
\item HDFS
\item Postgres
\end{itemize}
\end{frame}
%***********************************************************
\begin{frame}{Models}

\begin{columns}
\begin{column}{0.5\textwidth}
\begin{itemize}
\item kNN
\item Supervised learning
\item Unsupervised learning
\item MLE
\item Gradient Descent
\item Newton's Method
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\begin{itemize}
\item Linear Regression
\item Generalized linear models
\item Support vector machines
\item Expectation Maximization
\item Hidden Markov models
\item Mixture Models
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%***********************************************************
\begin{frame}{Concepts}

\begin{columns}
\begin{column}{0.5\textwidth}
\begin{itemize}
\item Data Science
\item Big Data
\item File types
\item BYOD
\item Regular expressions
\item Regularization
\item Distance measures
\item Norms
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\begin{itemize}
\item Experimental design
\item Model evaluation metrics
\item Overfitting
\item Outliers
\item Bias vs. variance
\item The Curse of Dimensionality
\item Dimensionality reduction
\end{itemize}
\end{column}
\end{columns}
\end{frame}

%***********************************************************
\begin{frame}{How did we do?}

\begin{itemize}
\item Exercises
\item Assignments
\item Labs
\item Lectures
\item[]
\item More? Less?
\item Shorter? Longer?
\item Theory vs. Practice
\item Pace?
\end{itemize}
\end{frame}

%***********************************************************
\begin{frame}{What did we do well?}

\begin{itemize}
\item Favorite?
\item Most useful?
\item TAs?
\item[]
\item More? Less?
\item Shorter? Longer?
\item Theory vs. Practice
\end{itemize}
\end{frame}

%***********************************************************
\begin{frame}{Preparation?}

\begin{itemize}
\item Prereqs
\begin{itemize}
\item COMP 140 / CAAM 210
\item STAT 310 / 315 
\item Changes?

\end{itemize}

\end{itemize}
\end{frame}


\end{document}

